Incentivizing Exploration with Heterogeneous Utilities

### 2017-11-09 paper.tex
did some edits of the paper

\begin{algorithm}
\caption{Algorithm: Upper Bound}
\label{Alg1}
\begin{algorithmic}
\STATE Set n = 1 to denote the round number;
\FOR{ $t = 1, 2, 3, \cdots$}{
	\IF{the myopic probability of pulling an arm is less than $n^{-1}$ and we haven’t pulled that arm in this round}
	\STATE{ we pay whatever it takes to incentivize that arm}\ELSE 
		\STATE {let the agents play myopically}\ENDIF
		\STATE $n = min_{i} N(i,t)$
}\ENDFOR

@@

\begin{lemma}
For $t\in[t_n,t_{n+1}]$, the expected incentive payment $E[c(t)]$ is bounded above by $\Delta + Nm \sqrt{\frac{2}{n\pi}}$. 
\end{lemma}

@@

Now we are ready to prove our first main result, Theorem~\ref{rst:budget}.

\begin{proof}
If $|u_{i}^{j}-u_{i,t}^{j}|\leq \lambda$ is true $\forall i$, $\forall j$, then we know for those $\theta\in \{\theta:\theta\cdot u_{B(\theta)}-\max_{j\neq B(\theta)}\{\theta \cdot u_{j}\}\geq 2m\lambda\}$, they will correctly identify their best arm. Thus we know, in the $n^{th}$ round, if $|u_{i}^{j}-u_{i,t}^{j}|\leq \frac{p^{-1}(\frac{p}{2})}{2m}$ $\forall i$ and $\forall j$, and $n^{-1}<p/2$, we do not need to incentivize any arms. In order to have $n^{-1}\leq \frac{p}{2}$, we need $n\geq \frac{2}{p}$. Denote $n_1=\max\{n_{0}, \frac{2}{p}\}$. Denote 
\begin{align}
n_{1}=\max\bigg\{50, \frac{23.04\sigma^4}{\lambda\left(p^{-1}(\frac{p}{2})\right)^4},\min_{n}\{n:n^{-1}<\frac{p}{2}\}\bigg\}. \nonumber
\end{align}
Denote $\delta_{0}=p^{-1}(\frac{p}{2})>0$ (because of Assumption~\ref{A1}).The cumulative payment is bounded above by:
\begin{align}
&E[C(T)] \nonumber \\
= & \sum_{t=1}^{\infty}E[c(t)\mathbbm{1}\{\text{need to offer payment to an arm}\}] \nonumber \\
\leq &\sum_{n=1}^{\infty}\sum_{t=1}^{\infty}E[c(t)\mathbbm{1}\{t_{n-1}\leq t\leq t_{n}: \text{myopic action occurs regret for agents with } \nonumber \\
&\theta\in \{\theta:\theta\cdot u_{B(\theta)}-\max_{j\neq B(\theta)}\{\theta \cdot u_{j}\}\geq p^{-1}(n^{-1})\}\}] \nonumber \\
\leq &\sum_{n=n_{1}}^{\infty}\sum_{t=1}^{\infty}\left(\Delta+Nm\sqrt{\frac{2}{n\pi}}\right)P(\text{myopic action occurs regret for agents with }\nonumber \\
&\theta\in \{\theta:\theta\cdot u_{B(\theta)}-\max_{j\neq B(\theta)}\{\theta \cdot u_{j}\}\geq \delta_{0}\}|t_{n-1}\leq t\leq t_{n})P(t_{n-1}\leq t\leq t_{n})+\sum_{n=1}^{n_1}N\left(\Delta+\sqrt{\frac{2}{n\pi}}\right) \nonumber \\
\leq & \sum_{n=n_{1}}^{\infty}24Nm \exp\left(\frac{-1.8 n\delta_{0}^2}{32m^2\sigma^2}\right)\times Nn\left(\Delta+Nm\sqrt{\frac{2}{n_1\pi}}\right)+\sum_{n=1}^{n_1}N\left(\Delta+\sqrt{\frac{2}{n\pi}}\right) \nonumber 
\end{align}

@@

The expected regret occurred when we incentivize the agents is calculated in Theorem~\ref{rst:budget}. Thus the cumulative regret at time $T$ is bounded above by
\begin{align}
&\sum_{n=1}^{n_{0}}NRn + \sum_{n=1}^{T} 24Nm\exp\left(-\frac{1.8n f(n)^2}{32 m^2\sigma^2}\right)\times R \times Nn+ \sum_{t=1}^{T}Mf(t)^2 \nonumber \\
+ & 24N^2 m \frac{e^{\frac{1.8\delta_{0}}{32m^2\sigma^2}}}{(e^{\frac{1.8\delta_{0}}{32m^2\sigma^2}}-1)^2}+N(\max\{n_{0},\frac{2}{p}\}). \nonumber
\end{align}

@@

\begin{algorithm}
\caption{Algorithm: Upper Bound}
\label{Alg2}
\begin{algorithmic}
\STATE Set n = 1 to denote the round number;
\FOR{ $t = 1, 2, 3, \cdots$}{
	\IF{the myopic probability of pulling an arm $i$ is less than $n^{-1}$ and we haven’t pulled that arm $i$ in this round (if there are multiple such arms, pick one randomly)}
	\STATE{Offer payment $c_{i,t}=\inf\{c: P(\theta\sim F: \theta\cdot u_{i,t}+c>max_{j}\theta\cdot u_{j,t})>n^{-1}\}$}\ELSE 
		\STATE {let the agents play myopically}\ENDIF
		\STATE $n = min_{i} N(i,t)$
}\ENDFOR

@@

For simplicity, we assume that the agents already know the true utility vector for arm $1$. Without loss of generality, denote $u_{t} = (0,1)+(z_{t1},z_{t,2}) = (0,1)+ (N(0,1/n),N(0,1/n))$ to be the estimate reward for arm $2$ (I assume the variance for the noise is $1$).

At time $n$, suppose we know arm $1$ has utility vector $(0,0)$ and arm $2$ has utility vector $u_n\sim (N(0,\frac{1}{n}), N(1,\frac{1}{n}))$. Since we know $(z_{t,1}, z_{t,2})$ is symmetric around $(0,0)$, we know

\begin{align}
&E[r_n] \nonumber \\
= &E[r_n | z_{n,1}>0,z_{n,2}>0] \times P(z_{n,1}>0,z_{n,2}>0) + E[r_n |z_{n,1}>0,z_{n,2}<0] \times P(z_{n,1}>0,z_{n,2}<0) \nonumber \\
= &E[r_n | z_{n,1}<0,z_{n,2}>0] \times P(z_{n,1}<0,z_{n,2}>0) + E[r_n |z_{n,1}<0,z_{n,2}<0] \times P(z_{n,1}<0,z_{n,2}<0) \nonumber \\
\geq & 0.25 \times E[r_n | z_{n,1}>0, z_{n,2}>0]. \nonumber
\end{align}

Given $z_{n,1}>0$ and $z_{n,2}>0$, we know the user between $(-1,0)$ and $(\frac{-1-z_{n,2}}{\sqrt{z_{n,1}^2+(1+z_{n,2})^2}}, \frac{z_{n,1}}{\sqrt{z_{n,1}^2+(1+z_{n,2})^2}})$ and the user between $(1,0)$ to $(\frac{1+z_{n,2}}{\sqrt{z_{n,1}^2+(1+z_{n,2})^2}}, \frac{-z_{n,1}}{\sqrt{z_{n,1}^2+(1+z_{n,2})^2}})$ will make a mistake. And the regret is the absolute value of the second coordinate of the user's preference vector. Thus, we know

\begin{align}
&E[r_n| z_{n,1}>0, z_{n,2}>0] \nonumber \\
=& 4\times 2 \int_{0}^{\infty} \int_{0}^{\infty} \int_{0}^{\arctan(\frac{z_{n,1}}{1+z_{n,2}})}\frac{\sin(\theta)}{2\pi}d(\theta)\frac{e^{-\frac{n \times z_{n,1}^2}{2}}\sqrt{n}}{\sqrt{2\pi}}d(z_{n,1})\frac{e^{-\frac{n \times z_{n,2}^2}{2}}\sqrt{n}}{\sqrt{2\pi}}d(z_{n,2}) \nonumber \\
=& \frac{2}{\pi^2}\int_{0}^{\infty} \int_{0}^{\infty}n\times [1-\frac{1+z_{n,2}}{\sqrt{z_{n,1}^2+(1+z_{n,2})^2}}]e^{-\frac{n \times z_{n,1}^2}{2}}e^{-\frac{n \times z_{n,2}^2}{2}}d(z_{n,1})d(z_{n,2}) \nonumber \\
=& \frac{2}{\pi^2}\int_{0}^{\infty} \int_{0}^{\infty} \left[1-\frac{\sqrt{n}+z_{n,2}}{\sqrt{z_{n,1}^2+(\sqrt{n}+z_{n,2})^2}}\right]e^{-\frac{z_{n,1}^2}{2}}e^{-\frac{z_{n,2}^2}{2}}d(z_{n,1})d(z_{n,2}) \nonumber 
\end{align}

Below, we want to show 
\begin{align}
\lim_{n\rightarrow\infty}\frac{E[r_n| z_{n,1}>0, z_{n,2}>0]}{n} = O(1). \nonumber
\end{align}

Denote $d(n)=n\left[1-\frac{\sqrt{n}+z_{n,2}}{\sqrt{z_{n,1}^2+(\sqrt{n}+z_{n,2})^2}}\right]$.
Since
\begin{align}
d^{'}(n)=\frac{-z_{n,1}^2(2z_{n,2}+3\sqrt{n})-2(z_{n,2}+\sqrt{n})^3}{2(z_{n,1}^2+(z_{n,2}+\sqrt{n})^2)^{3/2}}+1. \nonumber
\end{align}

and $\lim_{n\rightarrow \infty}d^{'}(n)=2$, we know for $n$ large enough, $d(n)$ is a increasing function in terms of n. Thus, based on the Monotone Convergence Theorem, we have

\begin{align}
&\lim_{n\rightarrow \infty}\frac{E[r_n| z_{n,1}>0, z_{n,2}>0]}{n} \nonumber \\
=& \frac{2}{\pi^2}\int_{0}^{\infty} \int_{0}^{\infty}\lim_{n\rightarrow \infty}\left[ \left[1-\frac{\sqrt{n}+z_{n,2}}{\sqrt{z_{n,1}^2+(\sqrt{n}+z_{n,2})^2}}\right]e^{-\frac{z_{n,1}^2}{2}}e^{-\frac{z_{n,2}^2}{2}}\right]d(z_{n,1})d(z_{n,2}) \nonumber  
\end{align}

@@

\begin{align}
\lim_{n\rightarrow \infty} d(n)=\frac{z_{n,1}^2}{2}. \nonumber
\end{align}
Thus,
\begin{align}
&\lim_{n\rightarrow \infty}\frac{E[r_n| z_{n,1}>0, z_{n,2}>0]}{n} \nonumber \\
=&\frac{2}{\pi^2}\int_{0}^{\infty} \int_{0}^{\infty}\lim_{n\rightarrow \infty}\left[ \frac{z_{n,1}^2}{2}e^{-\frac{z_{n,1}^2}{2}}e^{-\frac{z_{n,2}^2}{2}}\right]d(z_{n,1})d(z_{n,2}) = \frac{1}{2\pi}. \nonumber
\end{align}

>>>>

\begin{algorithm}
\caption{Algorithm: Incentivizing Exploration}
\label{Alg1}
\begin{algorithmic}
\STATE Set n = 1 to denote the round number;
\FOR{ $t = 1, 2, 3, \cdots$}{
	\IF{the myopic probability of pulling an arm is less than $n^{-1}$ and we haven’t pulled that arm in this round}
	\STATE{ we pay whatever it takes to incentivize that arm}\ELSE 
		\STATE {let the agents play myopically}\ENDIF
		\STATE $n = min_{i} N(i,t)$
}\ENDFOR

@@

\begin{lemma}
For $t\in[t_n,t_{n+1}]$, the expected incentive payment $E[c(t)]$ is bounded above by $\Delta + Nm \sqrt{\frac{2}{n\pi}}$. 
\label{lemma:payment}
\end{lemma}

@@

Now we are ready to prove our first main result, Theorem~\ref{rst:budget}. \bccomment{Not sure if I used Lemma~\ref{lemma:payment} correctly.}

\begin{proof}
If $|u_{i}^{j}-u_{i,t}^{j}|\leq \lambda$ is true $\forall i$, $\forall j$, then we know for those $\theta\in \{\theta:\theta\cdot u_{B(\theta)}-\max_{j\neq B(\theta)}\{\theta \cdot u_{j}\}\geq 2m\lambda\}$, they will correctly identify their best arm. Thus we know, in the $n^{th}$ round, if $|u_{i}^{j}-u_{i,t}^{j}|\leq \frac{p^{-1}(\frac{p}{2})}{2m}$ $\forall i$ and $\forall j$, and $n^{-1}\leq p/2$, we do not need to incentivize any arms. In order to have $n^{-1}\leq \frac{p}{2}$, we need $n\geq \frac{2}{p}$. Denote $n_1=\max\{n_{0}, \frac{2}{p}\}$. Denote $\delta_{0}=p^{-1}(\frac{p}{2})>0$ (because of Assumption~\ref{A1}).The cumulative payment is bounded above by:
\begin{align}
&E[C(T)] \nonumber \\
= & \sum_{t=1}^{\infty}E[c(t)\mathbbm{1}\{\text{need to offer payment to an arm}\}] \nonumber \\
\leq &\sum_{n=1}^{\infty}\sum_{t=1}^{\infty}E[c(t)\mathbbm{1}\{t_{n-1}\leq t\leq t_{n}: \text{myopic action occurs regret for agents with } \nonumber \\
&\theta\in \{\theta:\theta\cdot u_{B(\theta)}-\max_{j\neq B(\theta)}\{\theta \cdot u_{j}\}\geq p^{-1}(n^{-1})\}\}] \nonumber \\
\leq &\sum_{n=n_{1}}^{\infty}\sum_{t=1}^{\infty}\left(\Delta+Nm\sqrt{\frac{2}{n\pi}}\right)P(\text{myopic action occurs regret for agents with }\nonumber \\
&\theta\in \{\theta:\theta\cdot u_{B(\theta)}-\max_{j\neq B(\theta)}\{\theta \cdot u_{j}\}\geq \delta_{0}\}|t_{n-1}\leq t\leq t_{n})P(t_{n-1}\leq t\leq t_{n})+\sum_{n=1}^{n_1}N\left(\Delta+\sqrt{\frac{2}{n\pi}}\right) \nonumber \\
\leq & \sum_{n=n_{1}}^{\infty}24Nm \exp\left(\frac{-1.8 n\delta_{0}^2}{32m^2\sigma^2}\right)\times Nn\left(\Delta+Nm\sqrt{\frac{2}{n_1\pi}}\right)+\sum_{n=1}^{n_1}N\left(\Delta+\sqrt{\frac{2}{n\pi}}\right) \nonumber 
\end{align}

@@

The expected regret occurred when we incentivize the agents is calculated in Theorem~\ref{rst:budget}. Thus the cumulative regret at time $T$ is bounded above by
\begin{align}
&\sum_{n=1}^{n_{0}}NRn + \sum_{n=1}^{T} 24Nm\exp\left(-\frac{1.8n f(n)^2}{32 m^2\sigma^2}\right)\times R \times Nn+ \sum_{t=1}^{T}Mf(t)^2 \nonumber \\
+ & 24N^2 m \frac{e^{\frac{1.8\delta_{0}}{32m^2\sigma^2}}}{(e^{\frac{1.8\delta_{0}}{32m^2\sigma^2}}-1)^2}+N\left(\max\left\{n_{0},\frac{2}{p}\right\}\right). \nonumber
\end{align}

@@

\begin{algorithm}
\caption{Algorithm: Incentivizing Exploration Revised}
\label{Alg2}
\begin{algorithmic}
\STATE Set n = 1 to denote the round number;
\FOR{ $t = 1, 2, 3, \cdots$}{
	\IF{the myopic probability of pulling an arm $i$ is less than $n^{-1}$ and we haven’t pulled that arm $i$ in this round (if there are multiple such arms, pick one randomly)}
	\STATE{Offer payment $c_{i,t}=\inf\{c: P(\theta\sim F: \theta\cdot u_{i,t}+c>max_{j}\theta\cdot u_{j,t})>n^{-1}\}$}\ELSE 
		\STATE {let the agents play myopically}\ENDIF
		\STATE $n = min_{i} N(i,t)$
}\ENDFOR

@@

For simplicity, we assume that the agents already know the true utility vector for arm $1$. Without loss of generality, denote $u_{t} = (0,1)+(z_{t1},z_{t,2}) = (0,1)+ (N(0,1/t),N(0,1/t))$ to be the estimate reward for arm $2$ (Without loss of generality, we assume the variance for the noise is $1$).

At time $t$, suppose we know arm $1$ has utility vector $(0,0)$ and arm $2$ has utility vector $u_t\sim (N(0,\frac{1}{t}), N(1,\frac{1}{t}))$. Since we know $(z_{t,1}, z_{t,2})$ is symmetric around $(0,0)$, we know

\begin{align}
&E[r_t] \nonumber \\
= &E[r_t | z_{t,1}>0,z_{t,2}>0] \times P(z_{t,1}>0,z_{t,2}>0) + E[r_t |z_{t,1}>0,z_{t,2}<0] \times P(z_{t,1}>0,z_{t,2}<0) \nonumber \\
= &E[r_t | z_{t,1}<0,z_{t,2}>0] \times P(z_{t,1}<0,z_{n,2}>0) + E[r_t |z_{t,1}<0,z_{t,2}<0] \times P(z_{t,1}<0,z_{t,2}<0) \nonumber \\
\geq & 0.25 \times E[r_t | z_{t,1}>0, z_{t,2}>0]. \nonumber
\end{align}

Given $z_{t,1}>0$ and $z_{t,2}>0$, we know the user between $(-1,0)$ and $\left(\frac{-1-z_{t,2}}{\sqrt{z_{t,1}^2+(1+z_{t,2})^2}}, \frac{z_{t,1}}{\sqrt{z_{t,1}^2+(1+z_{t,2})^2}}\right)$ and the user between $(1,0)$ to $\left(\frac{1+z_{t,2}}{\sqrt{z_{t,1}^2+(1+z_{t,2})^2}}, \frac{-z_{t,1}}{\sqrt{z_{t,1}^2+(1+z_{t,2})^2}}\right)$ will make a mistake. And the regret is the absolute value of the second coordinate of the user's preference vector. Thus, we know

\begin{align}
&E[r_t| z_{t,1}>0, z_{t,2}>0] \nonumber \\
=& 4\times 2 \int_{0}^{\infty} \int_{0}^{\infty} \int_{0}^{\arctan\left(\frac{z_{t,1}}{1+z_{t,2}}\right)}\frac{\sin(\theta)}{2\pi}d(\theta)\frac{e^{-\frac{t \times z_{t,1}^2}{2}}\sqrt{t}}{\sqrt{2\pi}}d(z_{t,1})\frac{e^{-\frac{t \times z_{t,2}^2}{2}}\sqrt{t}}{\sqrt{2\pi}}d(z_{t,2}) \nonumber \\
=& \frac{2}{\pi^2}\int_{0}^{\infty} \int_{0}^{\infty}t\times \left[1-\frac{1+z_{t,2}}{\sqrt{z_{t,1}^2+(1+z_{t,2})^2}}\right]e^{-\frac{t \times z_{t,1}^2}{2}}e^{-\frac{t \times z_{t,2}^2}{2}}d(z_{t,1})d(z_{t,2}) \nonumber \\
=& \frac{2}{\pi^2}\int_{0}^{\infty} \int_{0}^{\infty} \left[1-\frac{\sqrt{t}+z_{t,2}}{\sqrt{z_{t,1}^2+(\sqrt{t}+z_{t,2})^2}}\right]e^{-\frac{z_{t,1}^2}{2}}e^{-\frac{z_{t,2}^2}{2}}d(z_{t,1})d(z_{t,2}) \nonumber 
\end{align}

Below, we want to show 
\begin{align}
\lim_{t\rightarrow\infty}\frac{E[r_t| z_{t,1}>0, z_{t,2}>0]}{t} = O(1). \nonumber
\end{align}

Denote $d(t)=t\left[1-\frac{\sqrt{t}+z_{t,2}}{\sqrt{z_{t,1}^2+(\sqrt{t}+z_{t,2})^2}}\right]$.
Since
\begin{align}
d^{'}(t)=\frac{-z_{t,1}^2(2z_{t,2}+3\sqrt{t})-2(z_{t,2}+\sqrt{t})^3}{2(z_{t,1}^2+(z_{t,2}+\sqrt{t})^2)^{3/2}}+1. \nonumber
\end{align}

and $\lim_{t\rightarrow \infty}d^{'}(t)=2$, we know for $n$ large enough, $d(t)$ is a increasing function in terms of t. Thus, based on the Monotone Convergence Theorem, we have

\begin{align}
&\lim_{t\rightarrow \infty}\frac{E[r_t| z_{t,1}>0, z_{t,2}>0]}{t} \nonumber \\
=& \frac{2}{\pi^2}\int_{0}^{\infty} \int_{0}^{\infty}\lim_{t\rightarrow \infty}\left[ \left[1-\frac{\sqrt{t}+z_{t,2}}{\sqrt{z_{t,1}^2+(\sqrt{t}+z_{t,2})^2}}\right]e^{-\frac{z_{t,1}^2}{2}}e^{-\frac{z_{t,2}^2}{2}}\right]d(z_{t,1})d(z_{t,2}) \nonumber  
\end{align}

@@

\begin{align}
\lim_{t\rightarrow \infty} d(t)=\frac{z_{t,1}^2}{2}. \nonumber
\end{align}
Thus,
\begin{align}
&\lim_{t\rightarrow \infty}\frac{E[r_t| z_{t,1}>0, z_{t,2}>0]}{t} \nonumber \\
=&\frac{2}{\pi^2}\int_{0}^{\infty} \int_{0}^{\infty}\lim_{t\rightarrow \infty}\left[ \frac{z_{t,1}^2}{2}e^{-\frac{z_{t,1}^2}{2}}e^{-\frac{z_{t,2}^2}{2}}\right]d(z_{t,1})d(z_{t,2}) = \frac{1}{2\pi}. \nonumber
\end{align}
